---
title: "Urgency Memo 4"
author: "Henrique Sposito"
date: "2024-03-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(poldis)
library(readr)
library(readxl)
```

# What are promises?

Promises in political discourse relate to future oriented  “calls to action” in relation to objects of policy. 

- only future (not past or present) - no taking stock but this activity can be a continuation of a previous action (e.g. "we will keep on working towards")
- "calls to action" are internally (e.g. "I will do") and externally (e.g. "we need to take this seriously") directed discourse aimed at describing, move, or inspiring future activity
- objects of policy refer to political problems constructed in discourse that demand dedicated policy attention - is this our subjects/topics?

Looking into promises allows us to rank political preferences related to future action.

```{r}
sample_text <- read_csv("sample_text.csv")
sentences <- annotate_text(sample_text$text, level = "sentences")
promises <- extract_promises(sentences)
subjects <- extract_subjects(promises)
related_terms <- extract_related_terms(promises$sentence, subjects)
urgency <- get_urgency(promises, related_terms)
head(urgency$sentence, n = 10)
```

# Outstanding issues:

- promise merging - maybe we can focus on sentence connectors for this (i.e. senetnecs that begin with these could be merged with previous sentences)?

```{r}
sentence_connectors <- read_excel("sentence_connectors.xlsx")
head(sentence_connectors, n = 20)
```

- normalization - bias towards longer sentences still an issue...
