---
title: "Is climate change a priority for US presidents?"
author: "Henrique Sposito, James Hollway, and Jael Tan"
date: '2026'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(poldis)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)
library(ggthemes)
library(messydates)
library(modelsummary)
```

# Data and methods

```{r}
# load and filter data
load("US_presidential_speeches_1993_2025.rda")
# get CC speeches only
US_clim_speeches <- US_speeches |>
  filter(grepl("climate change|global warming|paris agreement",
               text, ignore.case = TRUE)) |>
  mutate(doc_id = paste0("text", row_number()),
         speaker = factor(case_when(speaker == 'William J. Clinton' ~
                                          'Clinton',
                              speaker == 'George W. Bush' ~ 'Bush',
                              speaker == 'Barack Obama' ~ 'Obama',
                              speaker == 'Donald J. Trump' ~ 'Trump',
                              speaker == 'Joseph R. Biden, Jr.' ~ 'Biden'),
                              levels = c('Clinton', 'Bush', 'Obama',
                                         'Trump', 'Biden')))
# get priorities
cc_priorities <- select_priorities(US_clim_speeches$text)
#saveRDS(cc_priorities, file = "cc_priorities.rds")
# assign urgency scores
cc_urgency <- get_urgency(cc_priorities, summarise = "mean")
# filter for climate change, health, migration, and employment priorities
clim_urgency <- filter(cc_urgency,
                       grepl("climate change|global warming|paris agreement",
                             priorities, ignore.case = TRUE) &
                        Commitment > 0) |> # Remove non-priorities
  left_join(US_clim_speeches, by = "doc_id") |>
  mutate(cat = "Climate change")
```

# Urgency of climate change in American presidential speeches since 1993

```{r}
# plot CC by speaker
clim_urgency |>
  group_by(date, speaker) |>
  summarise(max_urgency = max(Urgency, na.rm = TRUE)) |>
  ggplot(aes(x = date, y = max_urgency)) +
  geom_point(aes(color = speaker), alpha = 0.4) +
  geom_smooth(aes(color = speaker), alpha = 0.25,
              method = "glm", se = TRUE, size = 2) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y",
               limits = c(as.Date("1993/01/20"), as.Date("2025/01/20"))) +
  scale_y_continuous(breaks = c(0,0.5,1,2,10)) + 
  theme_clean(base_size = 12, base_family = "serif") +
  labs(x = "Year", y = "Max Urgency Scores", color = "") +
  theme(legend.position = "bottom", 
        plot.caption = element_text(hjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

# Personal patterns in expressing urgency

It is also interesting to see how US presidents express climate change priorities. Rhetorical patterns or strategies differ in, for example, use of intensity or timing. This can be informative as a baseline. Or it can help better understand how urgency is expressed differently by individuals or over the course of a politician’s terms in office.

```{r}
hth_cc_urgency <- filter(cc_urgency, grepl("health|pandemic|infection", 
                                     priorities, ignore.case = TRUE) &
                        Commitment > 0) |> # Remove non-priorities
  left_join(US_clim_speeches, by = "doc_id") |>
  mutate(cat = "Health")
mig_cc_urgency <- filter(cc_urgency, grepl("migration|migrant",
  priorities, ignore.case = TRUE) & Commitment > 0) |> # Remove non-priorities
  left_join(US_clim_speeches, by = "doc_id") |>
  mutate(cat = "Migration")
emp_cc_urgency <- filter(cc_urgency, grepl("employment|job|wage",
  priorities, ignore.case = TRUE) & Commitment > 0) |> # Remove non-priorities
  left_join(US_clim_speeches, by = "doc_id") |>
  mutate(cat = "Employment")
tdata <- rbind(clim_urgency, emp_cc_urgency, hth_cc_urgency, mig_cc_urgency) |>
  group_by(date, cat, speaker) |>
  summarise(Commitment = max(Commitment, na.rm = TRUE),
            Intensity = max(Intensity, na.rm = TRUE),
            Timing = max(Timing, na.rm = TRUE),
            Frequency = max(Frequency, na.rm = TRUE)) |> 
  pivot_longer(cols = Commitment:Frequency,
               names_to = "type", values_to = "score") |>
  mutate(speaker = factor(speaker, levels = c(
           'Clinton', 'Bush', 'Obama', 'Trump', 'Biden'))) |>
  filter(score != 0)
modelsummary(list("Commitment" = lm(score ~ speaker,
                     data = subset(tdata, type == "Commitment")),
            "Timing" = lm(score ~ speaker,
               data = subset(tdata, type == "Timing")),
            "Frequency" = lm(score ~ speaker,
               data = subset(tdata, type == "Frequency")),
            "Intensity" = lm(score ~ speaker,
               data = subset(tdata, type == "Intensity"))),
            estimate = "{estimate}{stars}",
            coef_omit = "Intercept",
            coef_rename = c("speakerClinton" = "Clinton",
                            "speakerBush" = "Bush",
                            "speakerObama" = "Obama",
                            "speakerTrump" = "Trump",
                            "speakerBiden" = "Biden"))
```

The table above presents estimates obtained by fitting linear models to commitment, intensity, frequency, and timing scores for presidents (using Clinton as the reference category). In comparison to Clinton, we that both Bush commits less to climate change action. However, intensity, frequency, and timing scores do not seem correlated to president in statistically significant ways. That is, there appears to be no significant difference in the frequency in which presidents employ different urgency dimensions in their speech. \footnote{Future work using urgency analysis should investigate further differences in "oral styles" and how these can be used to set baselines by speaker and for how the usage of different urgency terms could be weighted according to these baselines.} 

# Comparing climate change priorities to employment, health, and migration in speeches that mention climate change

```{r}
rbind(clim_urgency, emp_cc_urgency, hth_cc_urgency, mig_cc_urgency) |>
  group_by(date, cat) |>
  summarise(max_urgency = max(Urgency, na.rm = TRUE)) |>
  ggplot(aes(x = date, y = max_urgency)) +
  geom_smooth(aes(color = cat), alpha = 0.6, 
              method = "loess", se = FALSE) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y",
                limits = c(as.Date("1993/01/20"), as.Date("2025/01/20"))) +
  scale_y_continuous(breaks = c(0,0.5,1,2,10)) + 
  annotate("rect", xmin = as.Date("2020/01/20"), xmax = as.Date("2023/05/01"),
           ymin = 0.1, ymax = 1.1, alpha=0.2, color="plum3", fill="plum1") +
  theme_clean(base_size = 12, base_family = "serif") +
  labs(x = "Year", y = "Max Urgency Scores", color = "") +
  theme(legend.position = "bottom", 
        plot.caption = element_text(hjust = 0.5),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

COVID sensitivity of the measurement approach and the sensibleness of the measures produced regarding their time pressure components 

# Comparing climate change priorities to employment, health, and migration in all speeches

```{r}
# health
US_h_speeches <- US_speeches |> # get priorities
  filter(grepl("health|pandemic|infection",
               text, ignore.case = TRUE)) |>
  mutate(doc_id = paste0("text", row_number()),
         speaker = factor(case_when(speaker == 'William J. Clinton' ~
                                          'Clinton',
                              speaker == 'George W. Bush' ~ 'Bush',
                              speaker == 'Barack Obama' ~ 'Obama',
                              speaker == 'Donald J. Trump' ~ 'Trump',
                              speaker == 'Joseph R. Biden, Jr.' ~ 'Biden'),
                              levels = c('Clinton', 'Bush', 'Obama',
                                         'Trump', 'Biden')))
# get priorities
h_priorities <- select_priorities(US_h_speeches$text)
# assign urgency scores
h_urgency <- get_urgency(h_priorities, summarise = "mean")
# filter for health only priorities
h_urgency <- filter(h_urgency, grepl("health|pandemic|infection", 
                                     h_priorities, ignore.case = TRUE) &
                        Commitment > 0) |> # Remove non-priorities
  left_join(US_h_speeches, by = "doc_id") |>
  mutate(cat = "Health")
# migration
US_m_speeches <- US_speeches |>
  filter(grepl("migration|migrant",
               text, ignore.case = TRUE)) |>
  mutate(doc_id = paste0("text", row_number()),
         speaker = factor(case_when(speaker == 'William J. Clinton' ~
                                          'Clinton',
                              speaker == 'George W. Bush' ~ 'Bush',
                              speaker == 'Barack Obama' ~ 'Obama',
                              speaker == 'Donald J. Trump' ~ 'Trump',
                              speaker == 'Joseph R. Biden, Jr.' ~ 'Biden'),
                              levels = c('Clinton', 'Bush', 'Obama',
                                         'Trump', 'Biden')))
# get priorities
m_priorities <- select_priorities(US_m_speeches$text)
# assign urgency scores
m_urgency <- get_urgency(m_priorities, summarise = "mean")
# filter for migration only priorities
mig_urgency <- filter(m_urgency, grepl("migration|migrant",
  priorities, ignore.case = TRUE) & Commitment > 0) |> # Remove non-priorities
  left_join(US_clim_speeches, by = "doc_id") |>
  mutate(cat = "Migration")
# employment
US_e_speeches <- US_speeches |>
  filter(grepl("employment|job|wage",
               text, ignore.case = TRUE)) |>
  mutate(doc_id = paste0("text", row_number()),
         speaker = factor(case_when(speaker == 'William J. Clinton' ~
                                          'Clinton',
                              speaker == 'George W. Bush' ~ 'Bush',
                              speaker == 'Barack Obama' ~ 'Obama',
                              speaker == 'Donald J. Trump' ~ 'Trump',
                              speaker == 'Joseph R. Biden, Jr.' ~ 'Biden'),
                              levels = c('Clinton', 'Bush', 'Obama',
                                         'Trump', 'Biden')))
# get priorities
e_priorities <- select_priorities(US_e_speeches$text)
# assign urgency scores
e_urgency <- get_urgency(e_priorities, summarise = "mean")
# filter for migration only priorities
e_urgency <- filter(e_urgency, grepl("employment|job|wage",
  priorities, ignore.case = TRUE) & Commitment > 0) |> # Remove non-priorities
  left_join(US_clim_speeches, by = "doc_id") |>
  mutate(cat = "Employment")
```

COVID sensitivity of the measurement approach and the sensibleness of the measures produced regarding their time pressure components 
